---
title: "Data 624 - Homework 3"
author: "John Kellogg"
date: "2/27/2021"
output: 
 html_document:
  code_folding: hide
  toc: true
  toc_float: 
    collapsed: true
    smooth_scroll: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
library(tidyverse)
library(fpp2)
library(seasonal)
```


## Question 2: 

**The 'plastics' data set consists of the monthly sales (in thousands) of product A for a plastics manufacturer for five years.**

```{r}
plastics
```


### 2a. 
  **Plot the time series of sales of product A. Can you identify seasonal fluctuations and/or a trend-cycle?**

```{r}
autoplot(plastics)+
  labs(title = "Plastics")+
  geom_smooth(method = "lm")+
  ylab("Sales")+
  xlab("Year")

ggseasonplot(plastics, polar=TRUE)

ggseasonplot(plastics)+
  labs(title = "Plastics")+
  ylab("Sales")+
  xlab("Year")
```
    
* The data shows a distinctive seasonality where the sales ramp up to the mid year and fall back down around Aug-Sept.
  * Jan is the lowest sales
* The data also shows a gradual growth trend 
  
### 2b. 
  **Use a classical multiplicative decomposition to calculate the trend-cycle and seasonal indices.**

```{r}
Q2b <- decompose(plastics, type = "multiplicative")
autoplot(Q2b)
```

* Not entirely comfortable with the remainder as there still seems to be a bit of a seasonal cycle left over.  
  * Would recommend using other methods to decompose. 

### 2c. 
  **Do the results support the graphical interpretation from part a?**

* Yes
  * The seasonal cycle seen in the graphs from question A are represented
  * The upwards trend is present

### 2d. 
  **Compute and plot the seasonally adjusted data.**

```{r}
autoplot(plastics, series = "Data")+
  autolayer(seasadj(Q2b), series = "Seasonally Adjusted")+
  autolayer(trendcycle(Q2b), series = "Trend")+
  labs(title = "Plastics", subtitle = "Seasonally Adjusted")
```


### 2e. 
  **Change one observation to be an outlier (e.g., add 500 to one observation), and recompute the seasonally adjusted data. What is the effect of the outlier?**
  
```{r}
#sample(1:60,1,replace = FALSE) 
#Sample chose #6
q2e <- plastics
q2e[6] <- q2e[6] + 500

outlier <- q2e %>%
  decompose(type = "multiplicative")

autoplot(q2e, series = "Data")+
  autolayer(seasadj(outlier), series = "Seasonally Adjusted")+
  autolayer(trendcycle(outlier), series = "Trend")+
  labs(title = "Plastics", subtitle = "Seasonally Adjusted with outlier")
```

#### Plastics with no outlier
```{r}
plastics %>% decompose(type = "multiplicative")%>%
  autoplot()
```

#### Plastics with outlier
```{r}
q2e %>% decompose(type = "multiplicative") %>%
  autoplot()
```

* The outlier creates a spike in the data 
* It also affected the trend line by creating a dip for the start of year 2
* It did NOT affect the seasonality
* For this, since the spike was so early in the data, it is not clearly shown in the remainder 

### 2f. 
  **Does it make any difference if the outlier is near the end rather than in the middle of the time series?**
  
```{r}
q2mid <- plastics
q2mid[30] <- q2mid[30] + 500

outlier <- q2mid %>%
  decompose(type = "multiplicative")

autoplot(q2mid, series = "Data")+
  autolayer(seasadj(outlier), series = "Seasonally Adjusted")+
  autolayer(trendcycle(outlier), series = "Trend")+
  labs(title = "Plastics", subtitle = "Seasonally Adjusted with outlier in data middle")
```

```{r}
q2end <- plastics
q2end[55] <- q2end[55] + 500

outlier <- q2end %>%
  decompose(type = "multiplicative")

autoplot(q2end, series = "Data")+
  autolayer(seasadj(outlier), series = "Seasonally Adjusted")+
  autolayer(trendcycle(outlier), series = "Trend")+
  labs(title = "Plastics", subtitle = "Seasonally Adjusted with outlier in data end")
```

#### Plastics with NO outlier
  
```{r}
plastics %>% decompose(type = "multiplicative")%>%
  autoplot()
```
  
#### Plastics with outlier in middle of data
  
```{r}
q2mid %>% decompose(type = "multiplicative") %>%
  autoplot()
```
  
#### Plastics with outlier at the end of the data
  
```{r}
q2end %>% decompose(type = "multiplicative") %>%
  autoplot()
```
  
* Trend line  
  * There is a slight bump in the trend line when the outlier is in the middle
  * When the outlier is at the end the drop off is sharper
* Remainder
  * When the outlier is in the middle there is a clear spike
  * When the outlier is toward the end, the moving average method drops most of the value.
* Seasonal
  * When the spike in in the middle, the seasonal graphs show a distinctive spike at the beginning of the cycle
  * When the spike is at the end, there is very little change
  
## Question 3: 
    
  **Recall your retail time series data (from Exercise 3 in Section 2.10). Decompose the series using X11. Does it reveal any outliers, or unusual features that you had not noticed previously?**
  
```{r}
retaildata <- readxl::read_excel("retail.xlsx", skip=1) 
myts <- ts(retaildata[,"A3349640L"], frequency=12, start=c(1982,4)) 

autoplot(myts)
ggseasonplot(myts, polar=TRUE)

myts %>% decompose(type = "multiplicative") %>%
  autoplot()
```
  
```{r}
myts_x11 <- seas(myts, x11="")
autoplot(myts_x11)
```

* I'm sure this process "should" show more outliers, however, the column chosen in Assignment 1 was so riddled with outliers already this did not show anything new.
