---
title: "Project 1 Part B"
author: "John Kellogg"
date: "3/28/2021"
output: 
 html_document:
  code_folding: hide
  toc: true
  toc_float: 
    collapsed: true
    smooth_scroll: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fpp2)
library(readxl)
library(knitr)
library(urca)
library(lubridate)
library(tidyverse)
library(grid)
library(gridExtra)
library(imputeTS)
```

## Part B - Forecasting Power
### Assignment Question
Part B consists of a simple dataset of residential power usage for January 1998 until December 2013.  Your assignment is to model these data and a monthly forecast for 2014.  The data is given in a single file. The variable ‘KWH’ is power consumption in Kilowatt hours, the rest is straight forward.
  
### Data Management
```{r atm_data, warning=FALSE}
data_raw_b <- read_excel ('./data/ResidentialCustomerForecastLoad-624.xlsx', col_names = TRUE)%>%
  rename(DATE = 'YYYY-MMM')%>%
  mutate(DATE=as_date(paste(DATE, '01', sep='-'), format='%Y-%b-%d', tz='UTC')) 
```

```{r}
summary(data_raw_b)
```

* We need to check for outliers as there is a large jump from the min to the 1st Q.
* There is a single NA value
  * k-nearest neighbor will be used again
```{r outliers}
boxplot(data_raw_b$KWH,
        ylab = "KWH")
out <- boxplot.stats(data_raw_b$KWH)$out
out
```

* We do have an outlier which needs to be adjusted.
* I will zero out the value and apply the same k-nearest neighbor as with the other NA value with a 5 level Weighted moving average..

```{r outlier_repair}
data_raw_b <- data_raw_b%>%
  mutate(KWH = na_if(KWH, out))
```

```{r na_repair}
data_raw_b$KWH <- na_ma(data_raw_b$KWH, k = 5, weighting = "simple")
```

```{r}
summary(data_raw_b)
```

* The data now looks clean and ready for transforms

### Data Transforms

```{r time_series}
pwr <- data_raw_b
tspwr <- ts(pwr$KWH, frequency = 12, start = c(1998,1))
```

```{r}
grid.arrange(nrow=2,
  ggseasonplot(tspwr),
  ggsubseriesplot(tspwr))

ggtsdisplay(tspwr)
```
  
#### Box-Cox Transform
```{r}
pwr_lambda <- BoxCox.lambda(tspwr)
pwrbc <- BoxCox(tspwr, lambda = pwr_lambda)

grid.arrange(nrow=2,
  ggseasonplot(pwrbc),
  ggsubseriesplot(pwrbc))

ggtsdisplay(pwrbc)
pwrbc %>% ur.kpss() %>% summary
```

```{r diff_check}
ndiffs(tspwr)
ndiffs(pwrbc)
```

* I am not seeing a major change in the data using a Box-Cox transform.  
* I do see both formats still require at least one diff.
* The Test Statistic is very high

#### No Box-Cox with a single Diff
```{r}
ggtsdisplay(diff(tspwr, 12), points = FALSE)
tspwr %>% diff %>% ur.kpss() %>% summary
```

#### Box-Cox with a single Diff
```{r}
ggtsdisplay(diff(pwrbc, 12), points = FALSE)
pwrbc %>% diff %>% ur.kpss() %>% summary
```

* I think my final choice is to use the non Box-Cox data set going forward.  I am not seeing enough justification to use the transform.
* The Lag of 12 does seem to be appropriate to the set.
 * An MA term seems appropriate as there is 2 sets of decreasing lags at multiple of 12 
 
```{r}
pwr_model <- Arima(tspwr, order = c(1,1,2), seasonal=c(1,1,2))
summary(pwr_model)
checkresiduals(pwr_model)
```

```{r}
auto_model <- auto.arima(tspwr, d=1, approximation = FALSE)
summary(auto_model)
checkresiduals(auto_model)
```
   
* My model seems very close to the auto model 
* For the sake of this project, I will continue with my model as there is not enough significant difference between the two to justify the swap.

```{r}
pwr_forecast <- forecast(pwr_model, h = 12)
autoplot(pwr_forecast)
```

```{r}
signif(pwr_forecast$mean/1000000,3)
```

```{r}
export <- pwr_forecast$mean
write.csv(export, "Power Forecasts.csv")
```
 