---
title: "Data 624 Assignment 7"
author: "John Kellogg"
date: "4/17/2021"
output: 
 html_document:
  code_folding: hide
  toc: true
  toc_float: 
    collapsed: true
    smooth_scroll: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fpp2)
library(readxl)
library(knitr)
library(urca)
library(pls)
library(caret)
library(tidyverse)
library(glmnet)
library(elasticnet)
library(knitr)
```

# Assignment 7:
In Kuhn and Johnson do problems 6.2 and 6.3. There are only two but they consist of many parts.  
Please submit a link to your Rpubs and submit the .rmd file as well.

## 6.2 
Developing a model to predict permeability (see Sect. 1.4) could save significant
resources for a pharmaceutical company, while at the same time more
rapidly identifying molecules that have a sufficient permeability to become a
drug:

### a.
Start R and use these commands to load the data:
```{r}
library(AppliedPredictiveModeling)
data(permeability)
```

The matrix fingerprints contains the 1,107 binary molecular predictors
for the 165 compounds, while permeability contains permeability
response.

### b.
The fingerprint predictors indicate the presence or absence of substructures
of a molecule and are often sparse meaning that relatively few of the
molecules contain each substructure. Filter out the predictors that have
low frequencies using the nearZeroVar function from the caret package.
How many predictors are left for modeling?

```{r}
fingerprints_adjusted <- fingerprints[,-nearZeroVar(fingerprints)]
print(paste("Non-Sparse predictors: ", ncol(fingerprints[,-nearZeroVar(fingerprints)])))
```


### c.
Split the data into a training and a test set, pre-process the data, and
tune a PLS model. How many latent variables are optimal and what is
the corresponding re-sampled estimate of R2?  

```{r}
set.seed(42)

df <- as.data.frame(fingerprints[, nearZeroVar(fingerprints)]) %>% mutate(permeability = permeability)

split_data <- createDataPartition(df$permeability, p=.75, list = FALSE)
x_train <- fingerprints_adjusted[split_data,]
x_test <- fingerprints_adjusted[-split_data,]
y_train <- permeability[split_data,]
y_test <- permeability[-split_data,]

train_df <- df[split_data,]
test_df <- df[-split_data,]

plsfit <- train(x=x_train, y=y_train,
  method = "pls",
  tuneGrid = expand.grid(ncomp = 1:10),
  trControl = trainControl(method = "cv", number = 10)
)

title <- paste("Training set RMSE minimized at",
               plsfit$bestTune$ncomp, 
               "Components")

plsfit_graph <- data.frame(plsfit$results$ncomp,plsfit$results$RMSE)

ggplot(plsfit_graph, aes(x=plsfit.results.ncomp, y=plsfit.results.RMSE))+
  geom_line()+
  geom_point()+
  xlab("Number of Variables")+
  ggtitle(title)

plsfit$results

plsfit$results %>%
  filter(ncomp == plsfit$bestTune$ncomp) %>%
  select(ncomp, RMSE, Rsquared)
```

* After splitting the data 75/25, the PLS model is tuned.
  * 10-fold cross validation
* Optimal number of features in this case is 7 explaining ~45% of the variance

### d.
Predict the response for the test set. What is the test set estimate of R2?
```{r}
training_data <- data.frame(x_train, y_train)

pls_model <- train(
  permeability ~ ., data = train_df, method = "pls",
  center = TRUE,
  trControl = trainControl("cv", number = 10),
  tuneLength = 25
)

result <- pls_model$result %>%
  filter(ncomp == pls_model$bestTune$ncomp) %>%
  select(ncomp, RMSE, Rsquared)

pls_results <- data.frame(
  Model = "PLS",
  RMSE = result$RMSE,
  Rsquared = result$Rsquared
)

result
```

* Test model performance got the following:
  * R squared 0.557
  * RMSE 11.183

### e.
Try building other models discussed in this chapter. Do any have better
predictive performance?

```{r pcr}
pcr_model <- train(
  permeability ~ ., data = train_df, method = "pcr",
  center = TRUE,
  trControl = trainControl("cv", number = 10),
  tuneLength = 25
)

pcr_pred <- predict(pcr_model, test_df)

pcr_results <- data.frame(
  Model = "PCR",
  RMSE = RMSE(pcr_pred, test_df$permeability),
  Rsquared = R2(pcr_pred, test_df$permeability)
)
```

```{r ridge}
r <- model.matrix(permeability ~ ., data = train_df)
r_test <- model.matrix(permeability ~ ., data = test_df)

ridge_cv <- cv.glmnet(r, train_df$permeability, alpha=0)
ridge_model <- glmnet(r, train_df$permeability, alpha = 0, lambda = ridge_cv$lambda.min)

ridge_pred <- as.vector(predict(ridge_model, r_test))

ridge_results <- data.frame(
  Model = "RIDGE",
  RMSE = RMSE(ridge_pred, test_df$permeability),
  Rsquared = R2(ridge_pred, test_df$permeability)
)
```

```{r elastic}
fractionTune = seq(.3,.8,.1)
lambdaTune = seq(.05, .3, .1)

elastic_model <- train(x = x_train, y = y_train, 
                  method = "enet", 
                  trControl = trainControl(method = "repeatedcv", repeats = 5), 
                  preProcess = c("center", "scale"), 
                  tuneGrid = expand.grid(lambda = lambdaTune,
                                         fraction = fractionTune)
                )

elastic_pred <- elastic_model%>%
  predict(x_test)

elastic_results <- data.frame(
  Model = "ElasticNet",
  RMSE = RMSE(elastic_pred, test_df$permeability),
  Rsquared = R2(elastic_pred, test_df$permeability)
 )
```

```{r}
results <- rbind(pcr_results, ridge_results)
results <- rbind(results, elastic_results)
results <- rbind(results, pls_results)

kable(results)
```



### f.
Would you recommend any of your models to replace the permeability
laboratory experiment?

* ElasticNet has the lowest RMSE and Highest R squared.
* It would be the better model

## 6.3
A chemical manufacturing process for a pharmaceutical product was
discussed in Sect. 1.4. In this problem, the objective is to understand the relationship
between biological measurements of the raw materials (predictors),
measurements of the manufacturing process (predictors), and the response of
product yield. Biological predictors cannot be changed but can be used to
assess the quality of the raw material before processing. On the other hand,
manufacturing process predictors can be changed in the manufacturing process.
Improving product yield by 1% will boost revenue by approximately
one hundred thousand dollars per batch:

### a.
Start R and use these commands to load the data:
```{r}
library(AppliedPredictiveModeling)
library(RANN)
data(ChemicalManufacturingProcess)

```
The matrix processPredictors contains the 57 predictors (12 describing
the input biological material and 45 describing the process predictors)
for the 176 manufacturing runs. yield contains the percent yield for each
run.

### b.
A small percentage of cells in the predictor set contain missing values. Use
an imputation function to fill in these missing values (e.g., see Sect. 3.8).

```{r}
knn_model <- preProcess(ChemicalManufacturingProcess, "knnImpute")
CMP_predict <- predict(knn_model, ChemicalManufacturingProcess)
```


### c.
Split the data into a training and a test set, pre-process the data, and
tune a model of your choice from this chapter. What is the optimal value
of the performance metric?


```{r}
df <- CMP_predict %>%
  select_at(vars(-one_of(nearZeroVar(.,names = TRUE))))

split_data <- createDataPartition(df$Yield, p=.75, list = FALSE)

train_df <- df[split_data,]
test_df <- df[-split_data,]

plsfit <- train(Yield ~.,
  data = train_df,
  method = "pls",
  tuneGrid = expand.grid(ncomp = 1:10),
  trControl = trainControl(method = "cv", number = 10),
  tuneLength = 25
)

title <- paste("Training set RMSE minimized at",
               plsfit$bestTune$ncomp, 
               "Components")

plsfit_graph <- data.frame(plsfit$results$ncomp,plsfit$results$RMSE)

ggplot(plsfit_graph, aes(x=plsfit.results.ncomp, y=plsfit.results.RMSE))+
  geom_line()+
  geom_point()+
  xlab("Number of Variables")+
  ggtitle(title)

plsfit$results

plsfit$results %>%
  filter(ncomp == plsfit$bestTune$ncomp) %>%
  select(ncomp, RMSE, Rsquared)
```

* After splitting the data 75/25, the PLS model is tuned.
  * 10-fold cross validation
* Optimal number of features in this case is 4 explaining ~60% of the variance 

### d.
Predict the response for the test set. What is the value of the performance
metric and how does this compare with the re-sampled performance metric
on the training set?

```{r}
pls_pred <- predict(plsfit, test_df)

pls_results <- data.frame(
  Model = "PLS",
  RMSE = RMSE(pls_pred, test_df$Yield),
  Rsquared = R2(pls_pred, test_df$Yield)
 )
pls_results
```

* PLS model has higher RMSE but lower R Squared.
* Personally, I say not much change.

### e.
Which predictors are most important in the model you have trained? Do
either the biological or process predictors dominate the list?

```{r}
pls_important <- varImp(plsfit)$importance %>%
  as.data.frame()%>%
  rownames_to_column("Variable")%>%
  filter(Overall >= 50) %>%
  arrange(desc(Overall)) %>%
  mutate(importance = row_number())

pls_important%>%
  mutate(Variable = fct_reorder(Variable, Overall))%>%
  ggplot(aes(x=Overall, y=Variable))+
  geom_bar(stat = "identity")
```

* Manufacturing dominates the top predictors

### f.
Explore the relationships between each of the top predictors and the response.
How could this information be helpful in improving yield in future
runs of the manufacturing process?

```{r}
pls_top <- varImp(plsfit, scale = FALSE)
toporder <- order(abs(pls_top$importance), decreasing = TRUE)
top <- rownames(pls_top$importance)[toporder[c(1:3)]] 

coef_order <- tibble(
  plsfit_coef_name = rownames(pls_top$importance),
  plsfit_coef = coef(plsfit$finalModel)
)

top15 <- coef_order %>%
  arrange(desc(plsfit_coef))%>%
  slice_head(n=15)

top15
```

```{r}
library(gridExtra)
p1 <- qplot(data = df, Yield, ManufacturingProcess32)
p2 <- qplot(data = df, Yield, ManufacturingProcess34)
p3 <- qplot(data = df, Yield, ManufacturingProcess09)
p4 <- qplot(data = df, Yield, ManufacturingProcess06)
p5 <- qplot(data = df, Yield, ManufacturingProcess04)
p6 <- qplot(data = df, Yield, BiologicalMaterial03)
p7 <- qplot(data = df, Yield, ManufacturingProcess33)
p8 <- qplot(data = df, Yield, ManufacturingProcess19)
p9 <- qplot(data = df, Yield, ManufacturingProcess39)
p10 <- qplot(data = df, Yield, ManufacturingProcess45)
p11 <- qplot(data = df, Yield, ManufacturingProcess11)
p12 <- qplot(data = df, Yield, ManufacturingProcess15)
p13 <- qplot(data = df, Yield, BiologicalMaterial06)
p14 <- qplot(data = df, Yield, ManufacturingProcess10)
p15 <- qplot(data = df, Yield, ManufacturingProcess44)

grid.arrange(p1, p2, p3, p4, nrow = 2, ncol=2)
grid.arrange(p5, p6, p7, p8, nrow = 2, ncol=2)
grid.arrange(p9, p10, p11, p12, nrow = 2, ncol=2)
grid.arrange(p13, p14, p15, nrow = 2, ncol=2)

```

* I hoped to see a noticeable change as the Coeff dropped.
* There is not enough trend to present as findings

```{r}
df_correlation <- df[,top15$plsfit_coef_name]
df_correlation <- cbind(df_correlation, df$Yield)
names(df_correlation)[16] <- "Yield"

df_corr <- cor(df_correlation)
corrplot::corrplot(df_corr,method = "circle", order = "alphabet", type = "lower", tl.srt = 45)

```

* At quick glance, I would recommend to stay away from Manufacturing Process 4.
  * It is in the top5 coeff values but has the highest negative correlation to yield
* Would Recommend Process 32
  * Top of the coeff values and has the highest positive correlation to yield

