---
title: "Data 624 Assignment 8"
author: "John Kellogg"
date: "4/24/2021"
output: 
 html_document:
  code_folding: hide
  toc: true
  toc_float: 
    collapsed: true
    smooth_scroll: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fpp2)
library(readxl)
library(knitr)
library(urca)
library(pls)
library(caret)
library(tidyverse)
library(glmnet)
library(elasticnet)
library(knitr)
```

## Assignment
Do problems 7.2 and 7.5 in Kuhn and Johnson. There are only two but they have many parts. Please submit both a link to your Rpubs and the .rmd file.

## 7.2
Friedman (1991) introduced several benchmark data sets create by simulation. One of these simulations used the following nonlinear equation to
create data: 

where the x values are random variables uniformly distributed between [0, 1] (there are also 5 other non-informative variables also created in the simulation). The package mlbench contains a function called mlbench.friedman1 that simulates these data:

```{r mlbench, warning=FALSE}
library(mlbench)
set.seed(200)
trainingData <- mlbench.friedman1(200, sd = 1)
## We convert the 'x' data from a matrix to a data frame
## One reason is that this will give the columns names.
trainingData$x <- data.frame(trainingData$x)
## Look at the data using
featurePlot(trainingData$x, trainingData$y)
## or other methods.
## This creates a list with a vector 'y' and a matrix
## of predictors 'x'. Also simulate a large test set to
## estimate the true error rate with good precision:
testData <- mlbench.friedman1(5000, sd = 1)
testData$x <- data.frame(testData$x)
```

Which models appear to give the best performance? Does MARS select the informative predictors (those named X1–X5)?

### KNN Model  
Tune several models on these data. For example:  

```{r knnmodel1, warning=FALSE}
library(caret)
knnModel <- train(x = trainingData$x,
 y = trainingData$y,
 method = "knn",
 preProc = c("center", "scale"),
 tuneLength = 10)

knnModel
```


```{r knnpred1, warning=FALSE}
knnPred <- predict(knnModel, newdata = testData$x)
## The function 'postResample' can be used to get the test set
## perforamnce values
postResample(pred = knnPred, obs = testData$y)
```

### MARS

```{r mars1, warning=FALSE}
mars_grid <- expand.grid(.degree = 1:2, .nprune = 2:15)
mars_model <- train(
  x = trainingData$x,
  y = trainingData$y,
  method = "earth",
  tuneGrid = mars_grid,
  preProcess = c("center","scale"),
  tuneLength = 10
        )

mars_model
```

```{r mars_pred, warning=FALSE}
marspredictions <- predict(mars_model, newdata = testData$x)
postResample(pred = marspredictions, obs = testData$y)
```

```{r mars_var, warning=FALSE}
varImp(mars_model)
```
  
### Neural Network
```{r NN1, warning=FALSE}
nngrid <- expand.grid(size = c(1:10),
                        decay = c(0, 0.01, 0.1),
                        bag = FALSE)

ctrl <- trainControl(method = "cv")

nnmodel <- train(trainingData$x, trainingData$y,
                  method = "avNNet",
                  tuneGrid = nngrid,
                  trControl = ctrl,
                  preProcess = c("center", "scale"),
                  linout = TRUE,
                  trace = FALSE,
                  MaxNWts = 10 * (ncol(trainingData$x) + 1) + 10 + 1,
                  maxit = 500
                  )

nnmodel
```

```{r nn_pred, warning=FALSE}
nnpredictions <- predict(nnmodel, newdata = testData$x)
postResample(pred = nnpredictions, obs = testData$y)
```

```{r nn_var, warning=FALSE}
varImp(nnmodel)
```
### SVM

```{r svm1, warning=FALSE}
svmmodel <- train(
  x = trainingData$x,
  y = trainingData$y,
  method = "svmRadial",
  preProcess = c("center","scale"),
  tuneLength = 10,
  trControl = trainControl(method = "cv")
        )
svmmodel
```

```{r svm_pred, warning=FALSE}
svmpredictions <- predict(svmmodel, newdata = testData$x)
postResample(pred = svmpredictions, obs = testData$y)
```

```{r svm_var, warning=FALSE}
varImp(svmmodel)
```

```{r results, warning=FALSE}
rbind(
  "mars" = postResample(pred = marspredictions, obs = testData$y),
  "svm" = postResample(pred = svmpredictions, obs = testData$y),
  "net" = postResample(pred = nnpredictions, obs = testData$y),
  "knn" = postResample(pred = knnPred, obs = testData$y)
)
```
  
* The Mars model has the losest RMSE, the highest R2, and the lowest MAE.  
* X1-X5 are the top predictors in all models ran.  
* Analysis: I would chose the MARS model unless some other reason presented itself.  
  
## 7.5
Exercise 6.3 describes data for a chemical manufacturing process. Use
the same data imputation, data splitting, and pre-processing steps as before
and train several nonlinear regression models.

### a.
Which nonlinear regression model gives the optimal resampling and test
set performance?  
  
```{r APM, warning=FALSE}
library(AppliedPredictiveModeling)
data("ChemicalManufacturingProcess")
set.seed(42)

preP <- preProcess(ChemicalManufacturingProcess, 
                   method = c("BoxCox", "knnImpute", "center", "scale"))
df <- predict(preP, ChemicalManufacturingProcess)

df$Yield = ChemicalManufacturingProcess$Yield

trainRows <- createDataPartition(df$Yield, p = .80, list = FALSE)
df.train <- df[trainRows, ]
df.test <- df[-trainRows, ]
```
  
```{r train,  warning=FALSE}
exclude <- which(colnames(df) == "Yield")
xtrain <- df.train[, -exclude]
ytrain <- df.train$Yield
xtest <- df.test[, -exclude]
ytest <- df.test$Yield
```

#### MARS
```{r MARS2, warning=FALSE}
marsGrid <- expand.grid(.degree=1:2,
                        .nprune=2:10)

mars_model <- train(
  x = xtrain,
  y = ytrain,
  method = "earth",
  preProcess = c("center","scale"),
  tuneGrid = marsGrid,
  tuneLength = 10
        )

mars_model
```

```{r mars_pred2, warning=FALSE}
marspredictions <- predict(mars_model, xtest)

mars <- postResample(pred = marspredictions, obs = ytest)
```
  
#### SVM
```{r svm2, warning=FALSE}
svmmodel <- train(
  x = xtrain,
  y = ytrain,
  method = "svmRadial",
  preProcess = c("center","scale"),
  tuneLength = 10,
  trControl = trainControl(method = "cv")
        )

svmmodel
```

```{r svm_pred2, warning=FALSE}
svmpredictions <- predict(svmmodel, xtest)

svm <- postResample(pred = svmpredictions, obs = ytest)
```
  
#### Neural Network
```{r nn2, warning=FALSE}
nngrid <- expand.grid(size = c(1:10),
                        decay = c(0, 0.01, 0.1),
                        bag = FALSE)

ctrl <- trainControl(method = "cv")

nnmodel <- train(xtrain, ytrain,
                  method = "avNNet",
                  tuneGrid = nngrid,
                  trControl = ctrl,
                  preProcess = c("center", "scale"),
                  linout = TRUE,
                  trace = FALSE,
                  maxit = 500
                  )

nnmodel
```

```{r nn_pred2, warning=FALSE}
nnpredictions <- predict(nnmodel, xtest)

nn <- postResample(pred = nnpredictions, obs = ytest)
```
  
#### KNN
```{r knn2, warning=FALSE}
knnModel <- train(x = xtrain,
 y = ytrain,
 method = "knn",
 preProc = c("center", "scale"),
 tuneLength = 10)

knnModel
```

```{r knn_pred2, warning=FALSE}
knnpredictions <- predict(knnModel, xtest)

knn <- postResample(pred = knnpredictions, obs = ytest)
```


```{r results2, warning=FALSE}
rbind(
  "mars" = mars,
  "svm" = svm,
  "net" = nn,
  "knn" = knn
)
```

* The MARS model performed better with the RMSE, however, its not going to be my choice.
* The SVM model has a slightly higher RMSE but lower r2 and lower Mean Absolute error.

### b.
Which predictors are most important in the optimal nonlinear regression
model? Do either the biological or process variables dominate the
list? How do the top ten important predictors compare to the top ten
predictors from the optimal linear model?

```{r svm_var2, warning=FALSE}
varImp(svmmodel, 10)
```

```{r svm_mars2, warning=FALSE}
varImp(mars_model, 10)
```

* Manufacturing Processes dominate the top 10 predictors
* Process 32 is the top predictor in both of the top models 

### c.
Explore the relationships between the top predictors and the response for
the predictors that are unique to the optimal nonlinear regression model.
Do these plots reveal intuition about the biological or process predictors
and their relationship with yield?

```{r graphics, warning=FALSE}
library(gridExtra)
P1 <- ggplot(df.train, aes(ManufacturingProcess32, Yield))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)

P2 <- ggplot(df.train, aes(ManufacturingProcess13, Yield))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)

P3 <- ggplot(df.train, aes(ManufacturingProcess09, Yield))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)

P4 <- ggplot(df.train, aes(ManufacturingProcess17, Yield))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)

P5 <- ggplot(df.train, aes(BiologicalMaterial06, Yield))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)

P6 <- ggplot(df.train, aes(BiologicalMaterial03, Yield))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)

grid.arrange(P1, P2, P3, P4, P5, P6, nrow = 3)

```

* In the top 5 predictors, there appears to be a positive relationship on the odd predictors (1,3,5) and a negative relationship on the even predictors (2,4)

